Manuscript_Github_Pchabaudi_calculateTPM

### 2021-09-01 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_calculateTPM.Rmd
Chabaudi use csv count data, and no SBP at all

```{r tpm, results = 'hide', message = FALSE}
# # Calculating the transcript-per-million 
# 
# We need to write a function to calculate the TPM from the count data we have generated. This function uses the 'chabaudi_transcript_lengths' object created in the chunk above and should generate errors if it detects discrepencies between the Geneids so that the wrong transcript length is not used.

tpm = function(count_data, lengths = chabaudi_transcript_lengths) {

  #Print an error if the id columns do not match
  # first if statement to test that both have a GeneID column
  if(any(grepl(colnames(lengths), 
               pattern = 'id', 
               ignore.case = T)) == T & any(grepl(colnames(count_data), 
                                                  pattern = 'Geneid', 
                                                  #This needs to be Geneid when some of the samples have 'id' in them
                                                  ignore.case = T)) == T){
    if(all(count_data[grep(colnames(count_data), 
                           pattern = 'Geneid', ignore.case = T)] == lengths[grep(colnames(lengths), 
                                                                             pattern = 'id', ignore.case = T)])) {
      print('Count data and length gene ids both match')
    } else {
      stop('Error: Count data and length gene ids DO NOT MATCH, please make sure they are ordered identically.')
    }
  } else {
    stop('Error: Count data and/or length table supplied do not have an id column.')
  }

  #Remove the id column from supplied count_data
  if(any(grepl(colnames(count_data), pattern = 'Geneid',ignore.case = T)) == T){
    counts_tpm = count_data[,-grep(colnames(count_data), pattern = 'Geneid', ignore.case = T)]
  } else{
    counts_tpm <- count_data}

  #Remove the id column from supplied lengths
  if(any(grepl(colnames(lengths), pattern = 'id', ignore.case = T)) == T){
    lengths.table = lengths[,-grep(colnames(lengths), pattern = 'id', ignore.case = T)]
  }else{
    lengths.table <- lengths}
  lengths.table = matrix(lengths.table)
  #This may make lengths.table a vector which can bring up some dimension problems in the following for loop

  #Now for the actual calculation
  rpm <- counts_tpm
  for (i in 1:dim(counts_tpm)[2]) {
    for (j in 1:dim(counts_tpm)[1]) {
      rpm[j,i] <- (counts_tpm[j,i]/(lengths.table[j]/1000))
    }
    scale.factor <- sum(rpm[,i])/(1E6)
    counts_tpm[,i] <- apply(rpm[,i,drop=F],1,function(x){x/scale.factor})
  }
  if(any(grepl(colnames(count_data), pattern = 'Geneid', ignore.case = T)) == T){
    counts_tpm = bind_cols(count_data[,grep(colnames(count_data), pattern = 'Geneid', ignore.case = T), drop = FALSE], 
                           counts_tpm)
  } 
  return(counts_tpm)
}
```

```{r function_mround}
mround <- function(x,base){ 
  base*ceiling(x/base) 
} 
```

```{r function_my_max}
my_max <- function(x) ifelse( !all(is.na(x)), max(x, na.rm=T), NA) #max robust to NAs
```

```{r function_ordering}
ordering <- function(to_order, order_vector){
  order = unlist(sapply(order_vector, 
                        function(x){unique(str_subset(to_order, 
                                                      pattern = paste(x)))}))
  order_missing = unique(to_order[!to_order %in% order])
  unique(c(order, order_missing), fromLast = TRUE)
}

@@

```{r read_tpm_rescaled_data}

chab_count_table <- read_xlsx('Manuscript_Frontiers_chabaudi_24h_counts.xlsx')

twentyfour_ct_comb <- tpm(chab_count_table, 
                          lengths = chabaudi_transcript_lengths)

@@

#Setting the re-scaled version as the default from now on
#Setting twentyfour_tpm_ct core dataset - removed pseudogene and re-scaled
twentyfour_ct_comb <- twentyfour_tpm_rescale_nopseudo

@@

twentyfour_rMT_avg <- melt(twentyfour_rMT, 
                           variable.name = 'sample',
                           value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, pattern = '.+(?=_[[:digit:]])')) %>% 
  group_by(Geneid, stage) %>% 
  summarise(mean_tpm = mean(tpm)) %>% 
  dcast(Geneid ~ stage)

# twentyfour_SBP <- select(twentyfour_ct_comb, 
#                          matches('Geneid|SBP')) %>% 
#   rename_with(~ str_extract(.x,pattern = '(?<=SBP_).+'),
#               matches('SBP'))
# 
# twentyfour_SBP_avg <- melt(twentyfour_SBP, 
#                            variable.name = 'sample',
#                            value.name = 'tpm') %>% 
#   mutate(stage = str_extract(sample, pattern = '.+(?=_[[:digit:]])')) %>% 
#   group_by(Geneid, stage) %>% 
#   summarise(mean_tpm = mean(tpm)) %>% 
#   dcast(Geneid ~ stage)

twentyfour_avgtpm <- list(
  'rMT' = list(
    tpm = twentyfour_rMT_avg,
    experiment = 'rMT'
  # ),
  # 'SBP' = list(
  #   tpm = twentyfour_SBP_avg,
  #   experiment = 'SBP'
  )
)

>>>>

```{r sourcing_functions}
source('../rfunctions/uniquify_fun.R')
source('../rfunctions/ordering_fun.R')
source('../rfunctions/tpm_fun.R')
source('../rfunctions/usefulFunctions_mround_my-max_firstup_fun.R')
```

```{r order_vec}

@@

```{r calc_tpm}

chab_count_table <- read.csv('Manuscript_Suppinfo_chabaudi_24h_counts.csv',
                             row.names = 1)

twentyfour_ct_comb <- tpm(chab_count_table, 
                          lengths = chabaudi_transcript_lengths)
```

```{r tpm_rescale}

@@

#Setting the re-scaled version as the default from now on
#Setting twentyfour_tpm_ct core dataset - removed pseudogene and re-scaled
twentyfour_ct_comb <- twentyfour_tpm_rescale_nopseudo
```

```{r}

@@

twentyfour_rMT_avg <- melt(twentyfour_ct_comb, 
                           variable.name = 'sample',
                           value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, pattern = '.+(?=_[[:digit:]])')) %>% 
  group_by(Geneid, stage) %>% 
  summarise(mean_tpm = mean(tpm)) %>% 
  dcast(Geneid ~ stage)

twentyfour_avgtpm <- list(
  'rMT' = list(
    tpm = twentyfour_rMT_avg,
    experiment = 'rMT'
)
